Uncovering underlying high-level musical content in the time domain \\ \small Leveraging self-supervised neural networks, inductive bias, and aural skills to learn deep audio embeddings with applications to boundary detection tasks

### 2023-07-07 discussion/discussion.tex
Updates from Overleaf

Whether our model effectively decodes underlying high-level musical content remains open to scientific investigation. It's plausible that our technique possesses significant potential for nearly all content-based MIR downstream tasks, given its intention to be both sound-agnostic and content-sensitive. Queries persist about whether such a general-purpose audio representation can mimic human hearing \cite{Li2023MERT:Training, Turian2022HEAR:Representations}, or if it can accurately decode high-level musical content. Such questions remain unresolved due to the current lack of evaluative measures. This represents a 'thorn in our side' that we intend to address in future research.

We argue whether factors such as the size of the representation layer and the size of the raw audio model might be insufficient \cite{verydeep}. Furthermore, we posit that with additional time to iterate repeatedly over the dataset, the loss function could decrease further with minimal risk of overfitting. \ref{fig:lossf}

While our results may not yet rival state-of-the-art benchmarks, they offer an encouraging starting point. They affirm we're on the right path and reveal the potential of this research direction. Importantly, we've seen that progress doesn't have to be expensive—though time and potential copyright issues are considerations.

>>>>

Whether our model effectively decodes underlying high-level musical content remains open to scientific investigation. It's plausible that our technique possesses significant potential for nearly all content-based MIR downstream tasks, given its intention to be both sound-agnostic and content-sensitive. Queries persist about whether such a general-purpose audio representation can mimic human hearing \cite{Li2023MERT:Training, Turian2022HEAR:Representations}, or if it can accurately decode high-level musical content. Such questions remain unresolved due to the current lack of evaluative measures.

This represents a 'thorn in our side' that we intend to address in future research.

We argue whether factors such as the size of the representation layer and the size of the model \cite{verydeep} might be insufficient. Furthermore, we posit that with additional time to iterate repeatedly over the dataset, the loss function could decrease further with minimal risk of overfitting. \ref{fig:lossf}

While our results may not yet rival state-of-the-art benchmarks, they affirm we're on the right path and reveal the potential of this research direction. Importantly, we've seen that progress doesn't have to be expensive—though time and potential copyright issues are considerations.

### 2023-07-07 evaluation/evaluation.tex
Updates from Overleaf

\chapter{Evaluation}

\section{The \textit{embeddiogram}, a deep audio feature representation}

The \textit{embeddiogram}, a deep audio feature representation, is derived by applying our pre-trained neural network to sliding windowed segments across the audio signal, generating a sequence of embedding vectors. These relatively low-dimensional embedding vectors collectively form a two-dimensional description of the audio signal's musical content.

Below is a detailed explanation of how we compute the \textit{embeddiogram} from a given audio signal of length $N$. This process comprises loading the audio data, slicing the audio data into windowed segments, processing each window using our pre-trained model to produce a vector per window/time frame, collecting and stacking these embeddings, and finally, normalizing the resulting matrix.

\begin{enumerate}
\item \textbf{Load the audio data}: The audio data is loaded into memory as a one-dimensional array of length $N$.

\item \textbf{Slice the audio data}: The audio data is segmented into overlapping windows. Each window contains $w$ samples, and consecutive windows are separated by a hop size $h$. This gives a total of $H$ windows, defined as:
\begin{equation}
H = 1 + \left\lfloor \frac{N - w}{h} \right\rfloor
\end{equation}
In the case where $\left( N - w \right) \mod h > 0$, we have $H += 1$ to account for the final, potentially smaller window.

We have conducted experiments using a window duration of 4 seconds (\(w = 4 \times \text{{sr (Hz)}}\)). As a general rule of thumb, this duration corresponds to two 4/4 bars of a piece of music with a tempo where the quarter note equals 120 BPM.

We find this to be a good starting compromise solution, allowing enough time span to capture musical content while not being so large that downsampling reduces the information to an unintelligible vector.

\item \textbf{Process each window}: Each window of audio data is processed independently, passed through the pre-trained neural network, and transformed into an embedding vector. Formally, for each window $w_i$ of audio data, we have:
\begin{equation}
\text{embedding}_i = \text{model}(w_i)
\end{equation}

\item \textbf{Collect the embeddings}: The embedding vectors are collected and stacked together. Each row represents a feature vector for a given time frame to form the \textit{embeddiogram}, denoted as $\text{e}$:
\begin{equation}
\text{e} = \begin{bmatrix} \text{embedding}_1 \\ \text{embedding}_2 \\ \vdots \\ \text{embedding}_H \end{bmatrix}
\end{equation}

\item \textbf{Normalize the \textit{embeddiogram \ref{fig:embeddiogram}}}: The \textit{embeddiogram} is normalized to have a minimum value of 0 and a maximum value of 1 per dimension. The normalization process is given by:
\begin{equation}
E'_{ij} = \frac{e_{ij} - \min(E)}{\max(E) - \min(E)}
\end{equation}
\end{enumerate}

\begin{figure}[ht]
  \centering
  \begin{minipage}[b]{1.0\linewidth}
    \centering
    \includegraphics[width=\linewidth]{figures/images/355embeddiogramnormalized.png}
    \caption[Embeddiogram. Track 355 (SALAMI dataset).]{Embeddiogram. Track 355 (SALAMI dataset).}
    \label{fig:embeddiogram}
  \end{minipage}

  \begin{minipage}[b]{1.0\linewidth}
    \centering
    \includegraphics[width=\linewidth]{figures/images/355novelty.png}
    \caption[Novelty curve + peaks. Track 355 (SALAMI dataset).]{Novelty curve and peak detection. Track 355 (SALAMI dataset).}
    \label{fig:novelty}
  \end{minipage}
\end{figure}

\begin{figure}[ht]
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/images/355recurrencematrix.png} % first figure itself
        \caption[Track 355 (SALAMI dataset) Self-similarity matrix]{Self-similarity matrix computation of the embeddiogram. Track 355 (SALAMI dataset).}
        \label{fig:SSM}
    \end{minipage}\hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/images/355lagmatrixgaussiansmoothing.png} % second figure itself
        \caption[Track 355 (SALAMI dataset) SSLM computation of the embeddiogram]{Self-similarity lag matrix. Track 355 (SALAMI dataset).}
        \label{fig:SSLM}
    \end{minipage}
    \vfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/images/355cumulativematrix.png} % third figure itself
        \caption[Track 355 (SALAMI dataset). Q-matrix]{Cumulative matrix computation of the embeddiogram. Track 355 (SALAMI dataset) .}
        \label{fig:Q}
    \end{minipage}\hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/images/355transitivebinarysimilaritymatrix.png} % fourth figure itself
        \caption[Track 355 (SALAMI dataset). Transitive Binary Similarity Matrix]{Transitive Binary Matrix computation of the embeddiogram. Track 355 (SALAMI dataset).}
        \label{fig:TBSM}
    \end{minipage}
\end{figure}

@@

These deep audio features can be a foundation for current state-of-the-art methods, which are expected to receive as input standard traditional audio features. Consequently, they can be processed and manipulated like conventional features as described by \cite{unsuperMSA} and displayed in \ref{fig:novelty}, \ref{fig:SSM}, \ref{fig:SSLM}, \ref{fig:Q}, and \ref{fig:TBSM}. Employing such embeddings in a traditional music segmentation algorithm can achieve state-of-the-art performance \cite{deepfeaturesegment}.

%%%%%%%%%%%%%%%%%%%%%%%%

While the usefulness of deep audio embeddings can be evaluated in countless downstream tasks \cite{Li2023MERT:Training, Kim2020OneStrategies}, we have chosen music boundary detection (also known as track segmentation), a subset of music structure analysis (MSA) for its popularity \cite{Smith2013ATask}, complexity, and product-compelling nature. This interdisciplinary field aims to understand the structure of music \cite{Nieto2020Audio-BasedApplications}. Due to subjectivity, ambiguity, and data scarcity, audio-based MSA faces non-solved challenges like boundary placement ambiguity and similarity quantification \cite{NietoPerceptualMusic}.

\section{Evaluation dataset and metrics}

The developed model has been evaluated on the SALAMI dataset \cite{Smith2011DESIGNANNOTATIONS}. The \textit{Structural Analysis of Large Amounts of Music Information} project aims to execute comprehensive structural analyses on a diverse spectrum of music. In contrast to traditional methods, SALAMI segments musical pieces into distinct sections, integrating perceptual, functional, and transcription analyses. Despite certain limitations, this novel approach offers a comprehensive and nuanced understanding of musical structure.

@@

On the other hand, Recall measures the proportion of accurately detected boundaries against all reference boundaries, indicating the completeness or sensitivity of the algorithm.

@@

Lastly, the F-measure offers a harmonized view of precision and recall. A widely adopted boundary detection metric, it compares predicted and ground truth boundaries, outputting a score between 0 and 1. This score is calculated as the harmonic mean of Precision and Recall, effectively considering both under-segmentation and over-segmentation. Given the potential inaccuracies inherent in human annotations and prediction errors, the F-measure permits minor deviations between predicted and actual boundaries. Its tolerance threshold can be adjusted, allowing a predicted boundary to be deemed correct if it is within a predefined window of a ground truth boundary \cite{NietoPerceptualMusic, Turnbull2007ABOOSTING}.

>>>>

\chapter{Datasets and Evaluation}

While the usefulness of similar self-supervised learned embeddings can be evaluated in countless downstream tasks \cite{Li2023MERT:Training, Kim2020OneStrategies}, we have chosen music boundary detection. Also known as track segmentation and commonly tackled with spectral analysis, it is a subset of music structure analysis (MSA) suitable for its popularity \cite{Smith2013ATask}, complexity, and product-compelling nature.

This interdisciplinary field aims to understand the structure of music \cite{Nieto2020Audio-BasedApplications}. Due to subjectivity, ambiguity, and data scarcity, audio-based MSA faces non-solved challenges like boundary placement ambiguity and similarity quantification \cite{NietoPerceptualMusic}.

@@

\section{Datasets and metrics}

The developed model has been trained on GTZAN \cite{GTZAN} and the Million Song Dataset (MSD) \cite{MSD} and evaluated on the SALAMI dataset \cite{Smith2011DESIGNANNOTATIONS}.

\subsection{GTZAN dataset}
The GTZAN Genre Collection is a widely-used dataset for music genre classification tasks.

This dataset comprises 1,000 audio tracks that are each 30 seconds long. These tracks are evenly distributed across ten genres: blues, classical, country, disco, hip hop, jazz, metal, pop, reggae, and rock, each containing precisely 100 songs. The audio files are stored as WAV files with a sample rate of 22050 Hz \cite{GTZAN}.

\subsection{Million Song Dataset (MSD)}
The Million Song Dataset (MSD) is a publicly available audio and metadata collection for a million contemporary popular music tracks.

The MSD encourages research on large-scale recommendation systems, exploration of musicological properties, and general research on large datasets. It provides a massive scale and diversity of data, making it an excellent resource for complex, large-scale music-related machine-learning tasks \cite{MSD}.

\subsection{SALAMI dataset}

The Structural Analysis of Large Amounts of Music Information (SALAMI) project conducts extensive structural analyses on a wide variety of music. SALAMI segments music pieces into distinct sections, integrating different analyses, including perceptual, functional, and transcriptional. While there are some limitations, this approach offers a nuanced and thorough understanding of musical structure.

SALAMI covers an extensive range of music genres and styles, including but not limited to classical, jazz, popular, and world music. These pieces originate from diverse sources, including Codaich, the Internet Archive's Live Music Archive, the RWC Music Database, and the Isophonics database.

Each piece of music in SALAMI is accompanied by detailed metadata such as title, artist, duration, names of the annotators, and the time taken for the annotation process. This metadata is provided in multiple formats, catering to each source database's needs.

While SALAMI does not directly distribute audio, it directs users to corresponding audio files on streaming platforms. This makes it a valuable resource for researchers working on music structure analysis, genre classification, music summarization, and other related fields \cite{Smith2011DESIGNANNOTATIONS}.

\subsection{Metrics}

@@

On the other hand, recall measures the proportion of accurately detected boundaries against all reference boundaries, indicating the completeness or sensitivity of the algorithm.

@@

Lastly, the F-measure provides a harmonized measure of precision and recall. As a widely adopted metric for boundary detection, it compares predicted boundaries with ground truth ones, yielding a score between 0 and 1. This score, calculated as the harmonic mean of Precision and Recall, effectively accounts for both under-segmentation and over-segmentation. Given the inherent inaccuracies in human annotations and prediction errors, the F-measure allows for minor deviations between predicted and actual boundaries. The F-measure can adjust the tolerance threshold, permitting a predicted boundary to be considered correct if it falls within a predefined window of a ground truth boundary \cite{NietoPerceptualMusic, Turnbull2007ABOOSTING}. In this work, we opted for an F-measure tolerance of 0.5 seconds, a decision largely guided by established norms in existing literature and music information retrieval research. This 0.5-second tolerance strikes a balance between precision and flexibility, demanding accurate boundary predictions while accommodating slight variations inevitable due to the subjective nature of music segmentation.

We computed the F-measure for each track and then calculated the average rather than aggregating all tracks into a single dataset and calculating the F-measure on this combined set.

### 2023-07-07 methods/subsections/implementation.tex
Updates from Overleaf

\begin{figure}
    \centering
    \includegraphics{figures/images/applsci-08-00150-g001.png}
    \caption[]{Caption}
    \label{fig:enter-label}
\end{figure}

@@

>>>>

\begin{figure}
    \centering
    \includegraphics{figures/images/applsci-08-00150-g001.png}
    \caption[SampleCNN]{Caption}
    \label{fig:samplecnn}
\end{figure}

\begin{figure}
    \centering
    \includegraphics{figures/images/applsci-08-00150-g002.png}
    \caption[SampleCNN]{Caption}
    \label{fig:filters}
\end{figure}

@@

\subsection{The \textit{embeddiogram}, a deep audio feature representation}

The \textit{embeddiogram}, a deep audio feature representation, is derived by applying our pre-trained neural network to sliding windowed segments across the audio signal, generating a sequence of embedding vectors. These relatively low-dimensional embedding vectors collectively form a two-dimensional description of the audio signal's musical content.

Below is a detailed explanation of how we compute the \textit{embeddiogram} from a given audio signal of length $N$. This process comprises loading the audio data, slicing the audio data into windowed segments, processing each window using our pre-trained model to produce a vector per window/time frame, collecting and stacking these embeddings, and finally, normalizing the resulting matrix.

\begin{enumerate}
\item \textbf{Load the audio data}: The audio data is loaded into memory as a one-dimensional array of length $N$.

\item \textbf{Slice the audio data}: The audio data is segmented into overlapping windows. Each window contains $w$ samples, and consecutive windows are separated by a hop size $h$. This gives a total of $H$ windows, defined as:
\begin{equation}
H = 1 + \left\lfloor \frac{N - w}{h} \right\rfloor
\end{equation}
In the case where $\left( N - w \right) \mod h > 0$, we have $H += 1$ to account for the final, potentially smaller window.

We have conducted experiments using a window duration of 4 seconds (\(w = 4 \times \text{{sampling rate}}\)). As a general rule of thumb, this duration corresponds to two 4/4 bars of a piece of music with a tempo where the quarter note equals 120 BPM.

We find this a good starting compromise solution, allowing enough time to capture musical content while not being so large that downsampling reduces the information to an unintelligible vector.

\item \textbf{Process each window}: Each window of audio data is processed independently, passed through the pre-trained neural network, and transformed into an embedding vector. Formally, for each window $w_i$ of audio data, we have:
\begin{equation}
\text{embedding}_i = \text{model}(w_i)
\end{equation}

\item \textbf{Collect the embeddings}: The embedding vectors are collected and stacked together. Each row represents a feature vector for a given time frame to form the \textit{embeddiogram}, denoted as $\text{e}$:
\begin{equation}
\text{e} = \begin{bmatrix} \text{embedding}_1 \\ \text{embedding}_2 \\ \vdots \\ \text{embedding}_H \end{bmatrix}
\end{equation}

\item \textbf{Normalize the \textit{embeddiogram \ref{fig:embeddiogram}}}: The \textit{embeddiogram} is normalized to have a minimum value of 0 and a maximum value of 1 per dimension. The normalization process is given by:
\begin{equation}
E'_{ij} = \frac{e_{ij} - \min(E)}{\max(E) - \min(E)}
\end{equation}
\end{enumerate}

\begin{figure}[ht]
  \centering
  \begin{minipage}[b]{1.0\linewidth}
    \centering
    \includegraphics[width=\linewidth]{figures/images/355embeddiogramnormalized.png}
    \caption[Embeddiogram. Track 355 (SALAMI dataset).]{Embeddiogram. Track 355 (SALAMI dataset).}
    \label{fig:embeddiogram}
  \end{minipage}

  \begin{minipage}[b]{1.0\linewidth}
    \centering
    \includegraphics[width=\linewidth]{figures/images/355novelty.png}
    \caption[Novelty curve + peaks. Track 355 (SALAMI dataset).]{Novelty curve and peak detection. Track 355 (SALAMI dataset).}
    \label{fig:novelty}
  \end{minipage}
\end{figure}

\begin{figure}[ht]
    \centering
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/images/355recurrencematrix.png} % first figure itself
        \caption[Track 355 (SALAMI dataset) Self-similarity matrix]{Self-similarity matrix computation of the embeddiogram. Track 355 (SALAMI dataset).}
        \label{fig:SSM}
    \end{minipage}\hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/images/355lagmatrixgaussiansmoothing.png} % second figure itself
        \caption[Track 355 (SALAMI dataset) SSLM computation of the embeddiogram]{Self-similarity lag matrix. Track 355 (SALAMI dataset).}
        \label{fig:SSLM}
    \end{minipage}
    \vfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/images/355cumulativematrix.png} % third figure itself
        \caption[Track 355 (SALAMI dataset). Q-matrix]{Cumulative matrix computation of the embeddiogram. Track 355 (SALAMI dataset) .}
        \label{fig:Q}
    \end{minipage}\hfill
    \begin{minipage}{0.45\textwidth}
        \centering
        \includegraphics[width=0.9\textwidth]{figures/images/355transitivebinarysimilaritymatrix.png} % fourth figure itself
        \caption[Track 355 (SALAMI dataset). Transitive Binary Similarity Matrix]{Transitive Binary Matrix computation of the embeddiogram. Track 355 (SALAMI dataset).}
        \label{fig:TBSM}
    \end{minipage}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%

These deep audio features can be a foundation for current state-of-the-art methods, which are expected to receive as input standard traditional audio features. Consequently, they can be processed and manipulated like conventional features as described by \cite{unsuperMSA} and displayed in \ref{fig:novelty}, \ref{fig:SSM}, \ref{fig:SSLM}, \ref{fig:Q}, and \ref{fig:TBSM}. Employing such embeddings in a traditional music segmentation algorithm can achieve state-of-the-art performance \cite{deepfeaturesegment}.

### 2023-07-07 results/subsections/tables.tex
Updates from Overleaf

As stated in the previous section, the evaluation results were obtained using standard and commonly used MIR tools, frameworks, and metrics. These include the MSAF \cite{MSAF} implementation for segmentation algorithms, the SALAMI dataset \cite{Smith2011DESIGNANNOTATIONS} for evaluation ground truth, and $mireval$ Python package \cite{RaffelMir_eval:METRICS} for metric computation.

The GitHub repository containing all the code needed to run the experiments can be found \href{https://github.com/oriolcolomefont/Master-Thesis.git}{HERE}. The results are displayed in tables \ref{tab:comparison_table} and \ref{ta:results}.

@@

\begin{table}[h]
\centering
\small
\begin{tabularx}{\textwidth}{>{\centering\arraybackslash}p{4.5cm}>{\centering\arraybackslash}X>{\centering\arraybackslash}X>{\centering\arraybackslash}X}
\toprule
\textbf{Training dataset [Ref]}  & \textbf{Precision} & \textbf{Recall} & \textbf{F-measure} \\
\midrule
\addlinespace
GTZAN \cite{GTZAN} & 0.228 & 0.171 & 0.185 \\
\addlinespace
MSD \cite{MSD} &  \textbf{0.333} &  \textbf{0.280} & \textbf{0.288} \\
\addlinespace
\bottomrule
\end{tabularx}
\caption[GTZAN-trained versus MSD-trained embeddiograms]{\small{Comparison of precision, recall, and f-measure for GTZAN-trained versus MSD-trained embeddiograms on SALAMI dataset computed using the Structural Feature (SF)\cite{sf} algorithm.}}
\label{tab:comparison_table}
\end{table}

@@

\begin{table}[ht]
\begin{threeparttable}
\centering
\small
\begin{tabularx}{\textwidth}{
  >{\centering\arraybackslash}p{4.5cm}
  >{\centering\arraybackslash}X
  >{\centering\arraybackslash}X
  >{\centering\arraybackslash}X}
\toprule
\thead{\textbf{Authors [Ref], Year}} & \thead{\textbf{Input}\tnote{1}} & \thead{\textbf{Method}} & \thead{\textbf{F-measure}} \\
\midrule
Turnbull et al. \cite{Turnbull2007ABOOSTING}, 2007 & MFCCs, chromas, spectrogram & Boosted Decision Stump  & 0.378 \\
\addlinespace
Kaiser et al. \cite{27}, 2012 & SSM & Novelty measure  & 0.286 \\
\addlinespace
Sargent et al. \cite{34}, 2011 & MFCCs, chromas & Viterbi  & 0.356 \\
\addlinespace
McFee \& Ellis \cite{20}, 2013 & MLS & Fisher’s Linear Discriminant  & 0.317 \\
\addlinespace
Nieto \& Bello \cite{28}, 2014 & MFCCs, chromas & Checkerboard-like kernel  & 0.299 \\
\addlinespace
Cannam et al. \cite{29}, 2015 & Timbre-type histograms & HMM  & 0.213 \\
\addlinespace
Nieto \cite{30}, 2016 & CQT Spectrogram & Linear Discriminant Analysis  & 0.299 \\
\addlinespace
Cannam et al. \cite{29}, 2017 & Timbre-type histograms & HMM  & 0.212 \\
\addlinespace
Ullrich et. al \cite{22}, 2014 & MLS & CNN  & 0.465 \\
\addlinespace
Grill \& Schlüter \cite{4}, 2015 & MLS, SSLMs & CNN  & 0.523 \\
\addlinespace
Grill \& Schlüter \cite{Grill2015MusicAnnotations}, 2015 & MLS, PCPs, SSLMs & CNN  & 0.508 \\
\addlinespace
Hadria \& Peeterss \cite{35}, 2017 & MLS, SSLMs & CNN  & 0.291 \\
\addlinespace
McCallum \cite{deepfeaturesegment}, 2019 & CQT & CNN  & \textbf{0.535} \\
\addlinespace
Ours, 2023 & Raw waveforms & CNN  & 0.288 \\
\bottomrule
\end{tabularx}
\caption[Baseline. State-of-the-art table.]{Previous studies' boundary detection f-measure results using unsupervised methods for a 0.5s time-window tolerance. Only the top-performing algorithm for each year on the SALAMI dataset is displayed. This table has been extended from \cite{Hernandez-Olivan2021MusicFeatures}.}
\label{tab:comparison_table}
\begin{tablenotes}\footnotesize
\item[1] \textbf{Legend:} SSM: Self-Similarity Matrix, MLS: Mel Spectrogram, MFCC: Mel-Frequency Cepstral Coefficient, CQT: Constant Q Transform, PCP: Pitch Class Profile, SSLM: Self-Similarity Lag Matrix.
\end{tablenotes}
\end{threeparttable}
\end{table}

@@

>>>>

The evaluation results were obtained using standard and commonly used MIR tools, frameworks, and metrics. These include the MSAF \cite{MSAF} implementation for segmentation algorithms, the SALAMI dataset \cite{Smith2011DESIGNANNOTATIONS} for evaluation ground truth, and $mireval$ Python package \cite{RaffelMir_eval:METRICS} for metric computation.

Table \ref{ta:results} provides a comparative analysis across four different features using three different segmentation algorithms. The results reveal that the performance of the algorithms varies depending on the type of feature. For instance, the CQT feature exhibits the highest precision (0.570), recall (0.339), and f-measure (0.353) when processed with the Foote algorithm. This table also proves that it is possible to achieve competitive perfor-
mance to traditional handcrafted signal processing methods by learning only from unlabeled audio files.

The visual representation of these results is also provided in Figure \ref{fig:boxplotmetrics} to better understand the distribution of the metrics for each feature.

Table \ref{tab:comparison_table} and Figure \ref{fig:enter-label} offer a comparison of the current study's boundary detection F-measure results with those of previous studies using unsupervised methods. The most accurate result was reported in 2019 by McCallum \cite{deepfeaturesegment} who used a CNN on CQT and achieved an F-measure of 0.535. The current study's approach using raw waveforms as input and CNN as the method yielded an F-measure of 0.288.

@@

\begin{table}[ht]
\centering
\small
\begin{tabularx}{\textwidth}{>{\centering\arraybackslash}p{4.5cm}>{\centering\arraybackslash}X>{\centering\arraybackslash}X>{\centering\arraybackslash}X}
\toprule
\textbf{Training dataset [Ref]}  & \textbf{Precision} & \textbf{Recall} & \textbf{F-measure} \\
\midrule
\addlinespace
GTZAN \cite{GTZAN} & 0.228 & 0.171 & 0.185 \\
\addlinespace
MSD \cite{MSD} &  \textbf{0.333} &  \textbf{0.280} & \textbf{0.288} \\
\addlinespace
\bottomrule
\end{tabularx}
\caption[GTZAN-trained versus MSD-trained embeddiograms]{\small{Comparison of precision, recall, and F-measure for GTZAN-trained versus MSD-trained embeddiograms on SALAMI dataset computed using the Structural Feature (SF)\cite{sf} algorithm.}}
\label{tab:comparison_table}
\end{table}

Table \ref{tab:comparison_table} compares the performance of the Structural Feature (SF) algorithm on the SALAMI dataset, utilizing two distinct sets of embeddiograms as input features. These embeddiograms were respectively generated with the neural network trained using the GTZAN and MSD datasets.

The results show a clear advantage for the SF algorithm when it is trained on the MSD dataset, as compared to the GTZAN dataset. Specifically, the precision, recall, and f-measure are all higher for the MSD-trained algorithm.

The precision, recall, and f-measure metrics in Table \ref{tab:comparison_table} evaluate the performance of the Structural Feature (SF) algorithm using GTZAN and MSD-trained embeddiograms. Precision indicates the proportion of correctly predicted boundaries. The MSD-trained algorithm outperforms the GTZAN-trained one, scoring 0.333 versus 0.228. Recall measures the algorithm's success in detecting all actual boundaries; again, the MSD-trained version shows better performance with a score of 0.280 against GTZAN's 0.171. F-measure, a metric that balances precision and recall, also favors the MSD-trained algorithm, with a score of 0.288 compared to 0.185 for the GTZAN-trained version.

@@

\begin{table}[ht]
\begin{threeparttable}
\centering
\small
\begin{tabularx}{\textwidth}{
  >{\centering\arraybackslash}p{4.5cm}
  >{\centering\arraybackslash}X
  >{\centering\arraybackslash}X
  >{\centering\arraybackslash}X}
\toprule
\thead{\textbf{Authors [Ref], Year}} & \thead{\textbf{Input}\tnote{1}} & \thead{\textbf{Method}} & \thead{\textbf{F-measure}} \\
\midrule
Turnbull et al. \cite{Turnbull2007ABOOSTING}, 2007 & MFCCs, chromas, spectrogram & Boosted Decision Stump  & 0.378 \\
\addlinespace
Kaiser et al. \cite{27}, 2012 & SSM & Novelty measure  & 0.286 \\
\addlinespace
Sargent et al. \cite{34}, 2011 & MFCCs, chromas & Viterbi  & 0.356 \\
\addlinespace
McFee \& Ellis \cite{20}, 2013 & MLS & Fisher’s Linear Discriminant  & 0.317 \\
\addlinespace
Nieto \& Bello \cite{28}, 2014 & MFCCs, chromas & Checkerboard-like kernel  & 0.299 \\
\addlinespace
Cannam et al. \cite{29}, 2015 & Timbre-type histograms & HMM  & 0.213 \\
\addlinespace
Nieto \cite{30}, 2016 & CQT Spectrogram & Linear Discriminant Analysis  & 0.299 \\
\addlinespace
Cannam et al. \cite{29}, 2017 & Timbre-type histograms & HMM  & 0.212 \\
\addlinespace
Ullrich et. al \cite{22}, 2014 & MLS & CNN  & 0.465 \\
\addlinespace
Grill \& Schlüter \cite{4}, 2015 & MLS, SSLMs & CNN  & 0.523 \\
\addlinespace
Grill \& Schlüter \cite{Grill2015MusicAnnotations}, 2015 & MLS, PCPs, SSLMs & CNN  & 0.508 \\
\addlinespace
Hadria \& Peeterss \cite{35}, 2017 & MLS, SSLMs & CNN  & 0.291 \\
\addlinespace
McCallum \cite{deepfeaturesegment}, 2019 & CQT & CNN  & \textbf{0.535} \\
\addlinespace
Ours, 2023 & Raw waveforms & CNN  & 0.288 \\
\bottomrule
\end{tabularx}
\caption[Baseline. State-of-the-art table.]{Previous studies' boundary detection f-measure results using unsupervised methods for a 0.5s time-window tolerance. Only the top-performing algorithm for each year on the SALAMI dataset is displayed. This table has been extended from \cite{Hernandez-Olivan2021MusicFeatures}.}
\label{tab:comparison_table}
\begin{tablenotes}\footnotesize
\item[1] \textbf{Legend:} SSM: Self-Similarity Matrix, MLS: Mel Spectrogram, MFCC: Mel-Frequency Cepstral Coefficient, CQT: Constant Q-Transform, PCP: Pitch Class Profile, SSLM: Self-Similarity Lag Matrix.
\end{tablenotes}
\end{threeparttable}
\end{table}

@@

The GitHub repository containing all the code needed to run the experiments can be found \href{https://github.com/oriolcolomefont/Master-Thesis.git}{HERE}.
